Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-09-15 paper/abstract.tex
Revise the introduction

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: they
either gracefully recover or safely shut down as requirements dictate.
The DARPA Cyber Assured Systems Engineering program is developing
tools for design, analysis, and verification that enable systems
engineers to design-in cyber-resiliency in a Model-Based Systems
Engineering (MBSE) environment.  This paper describes automated model
transformations that introduce high-assurance cyber-resiliency
components into a system; in particular, filters that guard against
malformed input, as well as monitors that guard against
spoofing and other malicious behavior.  A formal specification in the
form of a code contract defines each high-assurance component and is
used to verify that the component addresses system-level cyber
requirements.  Implementations for these high-assurance components
are directly synthesized from their code contracts, and are
automatically proven to preserve the exact meaning of the code
contract all the way down to the binary code level.  The model
transformations are integrated into the Open Source AADL 
Tool Environment (OSATE).  We further report on two case
studies, one by formal method experts and the other by non-experts,
applying model transformations to two different systems to improve
cyber resiliency using high-assurance filters and monitors.  One is a
UAV system utilizing the Air Force Research Laboratory's
OpenUxAS services for route planning.  The other is a tablet used for
situational awareness in a CH-47F Chinook helicopter application.
The case studies demonstrate that our MBSE approach is feasible in
real-world systems engineering.

>>>>

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: gracefully recover or safely shut down as requirements dictate.
\brfcs\ is a model-based systems engineering environment that extends the Open Source AADL Tool Environment to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
This manuscript describes a mechanism implemented by \brfcs\ to create, and insert, high-assurance components into a system for cyber-resiliency.
Such components can be filters that guard against malformed input or monitors that guard against spoofing and other malicious behavior.
A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts to verify the specified behavior.
Once tested, the code contract is used to verify that the added component addresses system-level cyber requirements.
Implementations for these
high-assurance components are directly synthesized from their code
contracts and are backed up by proofs showing that high-level
specifications map in a meaning-preserving way to code.
The manuscript further reports on a case study that used \brfcs\ to cyber-harden a UAV system by inserting high assurance components to isolate the open source Air Force Research Laboratory's OpenUxAS services for
route planning.  The case study
demonstrates that synthesizing implementations from code contracts is feasible in real-world systems
engineering.

### 2022-09-15 paper/intro.tex
Revise the introduction

\egm{Rewrite to make clear the contribution here: code contracts, tests, and synthesis through SPLAT. 
Present the \emph{unverified} path as is done in the introduction.
See if it is possible to present the verified path (ping Konrad) to state the various correctness guarantees.}

@@

Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.  Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

BriefCASE incorporates model-level cyber analysis tools (presently
GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
examine AADL models for potential vulnerabilities and suggest
cyber-security requirements to mitigate them.  A library of
architectural transforms guides the system engineer through automated
model transformations that modify the architecture to address these
requirements, possibly inserting new high-assurance components into
the system.  Implementations for the new components are synthesized
from formal specifications using
SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
(Semantic Properties for Language and Automata Theory).

Formal
verification that the transformed system model meets its requirements
is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
Environment).
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
\agr\ is a {\em compositional assume-guarantee} style model checker
for AADL models that attempts to prove properties about one layer of
an architecture using properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then
automatically generated using the High Assurance Modeling and Rapid
Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
desired, this code can be targeted to the formally verified seL4
secure microkernel~\cite{sel4-2009}.

%% DSH  This paragraph seems mostly duplicative of the paragraph above; attempted to merge the two.
%A novel aspect of the BriefCASE approach is the ability to automatically transform the model to insert high-assurance %components to satisfy requirements added to mitigate cyber vulnerabilities. In this work, the behavior of the system is %modeled by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result %of the cyber-vulnerability analysis with the new requirements for mitigation.  The requirements assume properties on the %input and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to %implement such requirements, AGREE fails to prove the system meets these new requirements.

Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components. These transformations not only change the
architecture of the model by adding in new components; they also
generate a formal specification in the form of a \emph{code contract} in the \agr\ language for their intended behavior. 
A code contract describes the computation of a high-assurance component and not just what it computes.
These code contracts
are sufficient for model checking to prove that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the \agr\ 
code contracts for the high-assurance components to \ckml, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This paper describes in detail the synthesis path
from code contracts to \ckml\ code, providing a formal framework in
which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries). Assuming that the execution schedule
of the deployed cyber-hardened system is as intended by the AADL
model, and that the HAMR-generated communication fabric delivers
messages between components as expected, the \agr\ model checking
results then hold for the deployed system, \ie, the system will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, supply-chain issues, and
identity verification; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are not yet dealt with in
our work.  Here, we do not report on the invention of a new type of
filter or monitor, in terms of capability; instead, our contribution
is in the automated synthesis of security-improving components from
formal specifications, and a means to show that the synthesis is
correct.

The approach is illustrated by a simple example in
\secref{sec:example}. \agr\ verification and specification are
presented in \secref{sec:agree} and \secref{agree-semantics}.
Code contracts are defined in \secref{sec:code-contracts} with a method
for testing code contracts in the \agr\ framework before synthesis given
\secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}. A
case study applying these transformations to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning is presented in
Section~\ref{sec:case-study}. Here the transforms add filters to guard
against malformed input and monitors to guard against ground station
spoofing and malicious flight plans from OpenUxAS. The case study
system is significantly more complex than the simple example and shows
the viability of the modeling approach to a full-scale industrial
design.

The \brfcs\ tools are open source and publicly
available \cite{fmide}, as is the motivating example \cite{repo}, and
the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
Videos demonstrating the use of the \brfcs\ tools to build the UAV
example presented in Section~\ref{sec:case-study} are also
available \cite{case}.

>>>>

@@

Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
They instead rely on process-oriented checklists and guidelines.
Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
One such tool is the Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
These are referred to as \emph{transforms} by \brfcs.
Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components.
\brfcs\ automates inserting the components into the system, and the designer writes a \emph{code contract} for each component to define its intended behavior.

A code contract is a formal specification in the Assume Guarantee Reasoning
Environment (\agr) language.
\agr is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
The code contract language is Turing complete allowing the designer to specify arbitrarily complex behavior.
These contracts are unit tested for correctness with \emph{test contracts}.
Test contracts define test scenarios to be implemented by the code contract under test. 
\agr\ proves whether or not that code contract implementation is correct enabling to designer to iteratively test the component behavior inside the \brfcs\ environment.
Once the behavior of the code contract is verified, \agr proves that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of the approach discussed in this manuscript is the synthesis of the code contracts for the high-assurance components to \ckml, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This manuscript describes in detail the synthesis path
from code contracts to \ckml\ code, providing a formal framework in
which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries). 
A key contribution here is that the code contract semantics are defined in such a way that the \agr\ verification 
results for the code contract hold for the deployed component, \ie, the component will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

The manuscript further details a 
case study applying these transformations with \brfcs\ to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning.
OpenUxAS, as an open source product, is considered \emph{untrusted}.
The UAV system is thus transformed to be resilient to malicious behavior that may arise in the untrusted component.
Here the transforms add filters to guard
against malformed input and monitors to guard against malicious flight plans from OpenUxAS. The case study
system is complex and shows
the viability of the approach in potential full-scale industrial
design.

The \brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, and supply-chain issues; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are not yet dealt with in
our work.  Here, we do not report on the invention of a new type of
high-assurance component in terms of capability; instead, our contribution
is in the automated synthesis of security-improving components from
formal specifications and a means to show that the synthesis is
correct.

The approach is illustrated by a simple example in
\secref{sec:example}. \agr\ verification and specification are
presented in \secref{sec:agree} and \secref{agree-semantics}.
Code contracts are defined in \secref{sec:code-contracts} with a method
for testing code contracts in the \agr\ framework before synthesis given
\secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}.

\begin{comment}
  BriefCASE incorporates model-level cyber analysis tools (presently
  GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
  examine AADL models for potential vulnerabilities and suggest
  cyber-security requirements to mitigate them.  A library of
  architectural transforms guides the system engineer through automated
  model transformations that modify the architecture to address these
  requirements, possibly inserting new high-assurance components into
  the system.  Implementations for the new components are synthesized
  from formal specifications using
  SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
  (Semantic Properties for Language and Automata Theory).

  Formal
  verification that the transformed system model meets its requirements
  is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
  Environment).
  %The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
  \agr\ is a {\em compositional assume-guarantee} style model checker
  for AADL models that attempts to prove properties about one layer of
  an architecture using properties allocated to its subcomponents.
  Cyber-resilient code implementing the verified model is then
  automatically generated using the High Assurance Modeling and Rapid
  Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
  desired, this code can be targeted to the formally verified seL4
  secure microkernel~\cite{sel4-2009}.
\end{comment}
